Investigation of Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles

### 2023-04-20 Latex files/Manuscript_rev3.tex
Updated Manuscript_rev3

\journal{Journal of Quantitative Spectroscopy \& Radiative Transfer}

@@

Since their invention in the late 1980s, optical tweezers have found application 
in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to
 testing the fundamental assumptions of quantum mechanics  \cite{yin2013large}, thanks 
mainly to the ability of the tweezer to transduce and detect forces on the order of a 
few pico newtons. Going beyond forces, further structural, dynamic and chemical 
characterisation of complex trapped entities could provide useful information, as 
demonstrated in   areas such as metrology \cite{arita2020coherent} and colloidal
 aggregation \cite{burns1990optical}. Spectroscopic techniques such as Raman scattering 
\cite{gupta2014raman} have been used for chemical characterisation of trapped objects, 
while dynamical characterisation has been demonstrated using data from the tweezer's 
Quadrant Photo Detector by following the centre-of-mass Brownian motion of the trapped 
entity  \cite{friedrich2012tuning} and measuring rotation  of the centre-of-mass 
\cite{yifat2021facile}.

The possibility of integrating an \textit{additional} light scattering detection scheme 
within an optical trapping setup was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998}, 
where a single-mode optical fibre was aligned to detect the scattered
light from a trapped bead and study its Brownian motion.  While this provided dynamical 
information, structural information about the trapped bead was precluded as the measurement 
was obtained only at a single angle.  In this work, we propose a scheme that expands on the 
technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of 
angles (Figure~\ref{fig:setup}), combined with a novel Bayesian inference-based analysis
 technique to enable interpretation of the resulting multi-angle data as well as 
optimisation to provide maximal information from the signal.

To demonstrate the analysis, we study a simple, illustrative example of a trapped entity 
more complex than a single sphere \textit{i.e.} a trapped asymmetric dimer.  
As a paradigmatic example of extracting structural/dynamic information from the 
resulting scattering data, we explore how to estimate the dimer's instantaneous 
orientation from the scattering signals using Bayesian inference, as well as how 
to optimise the analysis by implementing 'prior knowledge' to obtain the most reliable 
estimate. As an example of the relatively sparse literature on measuring orientations of
 complex trapped objects, ref.~\cite{raudsepp2022estimating} employs imaging to study the 
orientation of trapped dimers: scattering can give more quantitative and potentially more
 rapid time-resolved information, but only of course if the scattering signals can be 
interpreted. Here, we first train a neural network to effectively identify the mapping
 between scattering signals and dimer orientation, by calculating the scattering signal 
from a  simulated asymmetric dimer undergoing Brownian motion in an optical trap and 
mapping to the known instantaneous orientation of the simulated dimer. We then show 
how Bayesian inference can be used to optimise extraction of the true dimer orientation
 from the light scattering signals. Furthermore we demonstrate how the model can be
 fine tuned in situations where measurement uncertainty becomes significant.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
\label{sec:Method}

@@

We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
shaped sphere clusters. We simulate the motion of a dimer trapped in a highly focused Gaussian beam by calculating its diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}. For a dimer of arbitrary sized spheres the diffusion tensor - which describes the Cartesian forces and torques - is a 6 $\times$ 6 matrix:
\begin{align}
	\begin{pmatrix}
		F_x \\ F_y \\ F_z \\ T_x \\ T_y \\ T_z 
	\end{pmatrix}
	= \begin{pmatrix}
	a_1 &   0  &     0     &   0  & d_1 &     0 \\
	 0  &  a_1 &     0     & -d_1 &  0  &     0 \\
	 0  &   0  & a_1 + a_2 &   0  &  0  &     0 \\
	 0  & -d_1 &     0     &  b_1 &  0  &     0 \\
	d_1 &   0  &     0     &   0  & b_1 &     0 \\
	 0  &   0  &     0     &   0  &  0  & b_1 + b_2     
	\end{pmatrix}
	\begin{pmatrix}
		v_{k0} - V_k \\ v_{k0} - V_k \\ v_{k0} - V_k \\
		\omega_x - \Omega_x \\ \omega_y - \Omega_y \\ \omega_z - \Omega_z
	\end{pmatrix}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Orientation estimation from scattering measurements}
\label{sec:Bayes}

@@

Our goal is to determine the orientation of the trapped dimer based on the scattering data $I(\hat{n}, \ \theta_k)$. Rather than aim immediately for an exact estimate of the dimer's orientation, for the purposes of interpretation of the scattering and optimisation of the measurement setup it is more convenient to discretize the possible orientation space into a number of possible reference orientations, which we can then use as 'classification categories' in a neural network methodology to map scattering data to orientation (see below for further discussion).  Here we choose $\textit{n}_{ref} \ = \ 30$ reference orientations $\hat{\bf n}_{\alpha}$  evenly distributed on a unit sphere \cite{Rey2006} (Figure~\ref{fig:dimer}b) leading to a maximum nearest-neighbour spacing between two neighbouring reference orientations of 0.895 radians. Using MSTM we compute the raw intensities at each of the measurement angles that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the number and position of detection fibres is technically arbitrary there are several constraining factors that limit our ability to infer useful information from the trapped object, see Section~\ref{sec:Discussion} for a detailed breakdown of our choice of detection angles. The raw intensities are normalized according to:
\begin{align}
\label{eq:scale}
  y_k(\hat{\bf n}_\alpha)
  = 
  \frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
  {\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}
where the denominator is simply the standard deviation across the set of values $I(\hat{\bf n}, \theta_k)$. The reference orientations, raw intensities, and scaled signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.

Note that the collected scattering signals are not necessarily simply related to their associated reference orientations: as is well known from such examples of the inverse scattering problem, the mapping between the structural space of variables such as orientation, and the space of scattering signals may be highly complex. Nevertheless, at least where the uncertainty in signal measurements is low (see below), we can predict the orientation from the scattering by utilising computational techniques such as neural networks. We thus utilised the Python machine learning program \textit{scikit-learn} to build a neural network for identifying the dimer's orientation from its light scattering signal. The network was trained by generating a database of random orientation vectors, calculating the corresponding light scattering signals, and then using the network to estimate the probability of a given signal coming from a dimer in a given reference orientation. The network's loss function was evaluated and used to improve the estimation, the network being trained until the improvement in the loss function was less than 0.0001. 
%
Importantly, the estimation provided by the neural network can be improved further by accounting for any prior information we know about the dimer,  utilising Bayesian inference to update the neural network's estimation:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
To evaluate the above estimation of dimer orientation from scattering signal, we use a Brownian simulation of a dimer in the optical trap (Section~\ref{sec:brownian}) to compare estimated most probable reference orientation, derived from the dimer's scattering through Eq.~\eqref{eq:bayes}, with the dimer's known \emph{actual} orientation $\hat{\bf s}$. MSTM provides calculated light scattering from the simulated dimer $I(\hat{\bf s}, \theta)$ and we use \eqref{eq:scale} to obtain normalized values at each measurement angle $\theta_k$,  $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$, from which we obtain $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$. Because we know the actual orientation $\hat{s}$ we can measure the error in the model's estimate by comparing the reference orientation closest to $\hat{s}$, denoted as $\hat{\bf n}_{best}$, with the most probable predicted orientation from Eq.~\eqref{eq:bayes}. An ideal result would be one where the probability distribution is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$:
\begin{align}
	\label{eq:best}
	p_{best} = 
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}
In reality the distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation, leading to some level 'confidence' in orientation prediction, which can be quantified by calculating the Kullback-Leibler divergence $K_l$ between the two distributions:
\begin{align}
	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
	\right]
	\label{eq;kullback}
\end{align}
where a larger value of $K_l$ indicates that our model is less confident in its
prediction of the dimer's orientation. The divergence $K_l$ thus illustrates the 'spread'
in the estimated dimer orientation probability --- a distribution strongly peaked at 
some value would give us more confidence in that value than a near-uniform distribution 
where the scattering measurement could imply a wide range of possible orientations --- 
but it does not directly indicate our estimate's actual accuracy. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and  Discussion}
\label{sec:Discussion}\subsection{Testing the Model}
\label{sec:test}
Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer ($n = 1.59$) trapped in water ($n = 1.33$) within a 5 mW optical trap. The trapping laser is 1064nm NIR focused through a 1.25 NA objective. The dimer is comprised of two tangent spheres with radii $1 \mu m$ and $0.5 \mu m$ respectively. We simulated the first 10 seconds of motion, calculating the orientation and position every 1 ms.

@@

One reason we observe such large jumps in  orientation estimated from scattering signals is that there is no simple correlation between the 'distance in scattering space' between scattering signals from two different orientations, and their separation in orientation space : even a large change in orientation can involve a small change in scattering. Combining this fact with use of a uniform prior, indicating essentially no knowledge of how orientation should behave, there is no constraint on how much estimated orientation can change from time-step to time-step. To improve the estimation we can therefore use knowledge of the physical limitations of the object in the trap and its dynamics, imposing a more physically grounded prior,  accounting in this case for the fact that the motion of the dimer is limited due to the trap stiffness. Here the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at each time step according to the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$, \textit{i.e.} taking into account the different likelihoods of the estimated orientation jumping between nearby or distant reference orientations:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Asymmetric dimer dynamics}
\label{sec:motion}

The Brownian OT software was used to simulate the motion of a trapped
dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) over the first
$10$ seconds of entering the optical trap.  The initial orientation
was assumed as strictly vertical (in line with the beam propagation
direction). The dimer's position and orientation was recorded every
$1\,{\rm ms}$ for using as a test dataset for our model.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{./Images/fig5a.png}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig5b.png}
	\end{subfigure}
	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
	\label{fig:motion}
\end{figure}

As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.  Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.  It is interesting to note that the dimer is furthest from the trap centre as it reaches a horizontal orientation before drawing closer again as it inverts completely. Further simulations of dimers with different size ratio $a_1 \geq 2a_2$ showed a similar behaviour of initially moving away from the trap focus and rotating while remaining trapped before re-aligning, while dimers with size ratis $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
%
In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque is minimal for a dimer in both horizontal and vertical orientations.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Impact of measurement noise on model predictions}
\label{sec:epsilon}

So far a key assumption of the neural network implementation is that the detected scattering signal  has no uncertainty associated with it. In reality of course detected scattering signals  will have some non-zero level of measurement  noise. To explore the impact of measurement uncertainty on orientation estimation model performance we introduce a Gaussian noise to the measured signal:
\begin{align}
	I(\hat{\bf s}) = I(\hat{\bf s}) \pm \epsilon I(\hat{\bf s})
\end{align}
where $\epsilon$ is the percentage error associated with the scattering signal.  Figure~\ref{fig:epsilon} shows the performance of the model at a range of $\epsilon$ using in-plane detector angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and out-of-plane detector at $75^{\circ}$, with $\beta$ set to $1$:

>>>>

\journal{}

@@

Since their invention in the late 1980s, optical tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to testing the fundamental assumptions of quantum mechanics \cite{yin2013large}, thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few pico-newtons. Going beyond forces, further structural, dynamic and chemical characterisation of complex trapped entities could provide useful information, as demonstrated in areas such as metrology \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}.  Spectroscopic techniques such as Raman scattering \cite{gupta2014raman} have been used for chemical characterisation of trapped objects, while dynamical characterisation has been demonstrated using data from tweezer's Quadrant Photo Detector (QPD) by following the centre-of-mass Brownian motion of the trapped entity \cite{friedrich2012tuning} and measuring rotation of the centre-of-mass\cite{yifat2021facile}.  A recent work aimed at characterising trapped entities demonstrated how neural networks maybe trained to distinguish between optically trapped micro-beads of different size and material by means of a principal component analysis of the forward scattered light detected using a QPD \cite{Carvalho_2023}.  A more direct, albeit cumbersome attempt at detecting scattered light from trapped biological cells was attempted in \cite{Watson_2023} where the experimental cuvette was placed inside an elliptical mirror that directed light scattered from the trapped biological-cell onto a photodetector via a rotating aperture that helped select the scattering-angle. Thus, past studies on optical trapping have focused rather heavily on particle-tracking but not particle-characterisation.

While both \cite{Carvalho_2023} \cite{Watson_2023} demonstrate some ability to characterise trapped entities, \cite{Carvalho_2023} is perhaps best suited to characterise micron-sized particles and \cite{Watson_2023} talks about an experimental setup that is difficult to adopt and suffers from a low-bandwidth that might not be best suited for monitoring dynamics. A light scattering detection scheme built around an optical trap that is both easy to implement and has the advantage of high-bandwidth was demonstrated by Safran and co-workers in \cite{Bar-Ziv_1998}, where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and study its Brownian motion. While this provided dynamical information, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle. Furthermore, the main drawback to a single fibre detection scheme is that it struggles to characterise isotropically scattering particles. This limits its application to biological matter as the internal structure of many cells makes them inherently isotropic scatterers \cite{Watson_2023}. In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at a number of angles (Figure 1), combined with a novel Bayesian inference-based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation to provide maximal information from the signal.

To demonstrate the analysis, we study a simple, illustrative example of an isotropic scattering entity \emph{i.e.} an asymmetric dimer. As a paradigmatic example of extracting structural/dynamic information from the resulting scattering data, we explore how to estimate the dimer's instantaneous orientation from the scattering signals using Bayesian inference, as well as how to optimise the analysis by implementing 'prior knowledge' to obtain the most reliable estimate. As an example of the relatively sparse literature on measuring orientations of complex trapped objects, ref. employs imaging to study the orientation of trapped dimers: scattering can give more quantitative and potentially more rapid time-resolved information, but only of course if the scattering signals can be interpreted. Here, we first train a neural network to effectively identify the mapping between scattering signals and dimer orientation, by calculating the scattering signal from a simulated asymmetric dimer undergoing Brownian motion in an optical trap and mapping to the known instantaneous orientation of the simulated dimer. We then show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals. Furthermore we demonstrate how the model can be fine tuned in situations where measurement uncertainty becomes significant.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}
\label{sec:Method}

@@

We use the Brownian OT package developed by Fung~\textit{et~al} \cite{Vigilante2020Brownian_OT} to simulate the motion of an asymmetric dimer (Figure~\ref{fig:dimer}a) within an optical trap. Brownian OT combines MSTM \cite{Mishchenko1996MSTM} and ``Optical Tweezer Toolbox'' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary
shaped sphere clusters. We simulate the motion of a dimer trapped in a 
highly focused Gaussian beam by calculating the optical forces imparted by 
the laser, and the Brownian forces imparted by the surrounding fluid. MSTM provides the necessary T-matrix to compute the optical force via \textit{ott}. The Brownian force is found by computing the dimer's diffusion tensor according to the analytical solutions provided by Nir and Acrivos \cite{nir_acrivos_1973}. For a dimer of arbitrary sized spheres the diffusion tensor - which describes the Cartesian forces and torques - is a 6 $\times$ 6 matrix:
\begin{align}
	\begin{pmatrix}
		F_x \\ F_y \\ F_z \\ T_x \\ T_y \\ T_z 
	\end{pmatrix}
	= \begin{pmatrix}
	a_1 &   0  &     0     &   0  & d_1 &     0 \\
	 0  &  a_1 &     0     & -d_1 &  0  &     0 \\
	 0  &   0  & a_1 + a_2 &   0  &  0  &     0 \\
	 0  & -d_1 &     0     &  b_1 &  0  &     0 \\
	d_1 &   0  &     0     &   0  & b_1 &     0 \\
	 0  &   0  &     0     &   0  &  0  & b_1 + b_2     
	\end{pmatrix}
	\begin{pmatrix}
		v_{k0} - V_k \\ v_{k0} - V_k \\ v_{k0} - V_k \\
		\omega_x - \Omega_x \\ \omega_y - \Omega_y \\ \omega_z - \Omega_z
	\end{pmatrix}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Orientation estimation from scattering measurements}
\label{sec:Bayes}

@@

Our goal is to determine the orientation of the trapped dimer based on the scattering data $I(\hat{n}, \ \theta_k)$. Rather than aim immediately for an exact estimate of the dimer's orientation, for the purposes of interpretation of the scattering and optimisation of the measurement setup it is more convenient to discretize the possible orientation space into a number of possible reference orientations, which we can then use as 'classification categories' in a neural network methodology to map scattering data to orientation (see below for further discussion).  Here we choose $\textit{n}_{ref} \ = \ 30$ reference orientations $\hat{\bf n}_{\alpha}$  evenly distributed on a unit sphere \cite{Rey2006} (Figure~\ref{fig:dimer}b) leading to a maximum nearest-neighbour spacing between two neighbouring reference orientations of 0.895 radians. Using MSTM we compute the raw intensities at each of the measurement angles that would be generated by a dimer in each reference orientation, $I(\hat{\bf n}_{\alpha}, \theta_k)$. While the number and position of detection fibres is technically arbitrary there are several constraining factors that limit our ability to infer useful information from the trapped object, see Section~\ref{sec:detectors} for a detailed breakdown of our choice of detection angles. The raw intensities are normalized according to:
\begin{align}
\label{eq:scale}
  y_k(\hat{\bf n}_\alpha)
  = 
  \frac{I(\hat{\bf n}_\alpha, \theta_k) - \langle I(\hat{\bf n},\theta_k) \rangle } 
  {\langle I^2(\hat{\bf n},\theta_k) \rangle -\langle I(\hat{\bf n}, \theta_k)\rangle^2}
\end{align}
where the denominator is simply the standard deviation across the set of values $I(\hat{\bf n}, \theta_k)$. The reference orientations, raw intensities, and scaled signals are given in Tables~\ref{tab:A1} and \ref{tab:A2}.

Note that the collected scattering signals are not necessarily simply related to their associated reference orientations: as is well known from such examples of the inverse scattering problem, the mapping between the structural space of variables such as orientation, and the space of scattering signals may be highly complex. Even if the orientation space is divided evenly between reference orientation the subsequent signal space ends up being appearing mixed making simple comparisons of signals useless for inferring information on the particle. 
Nevertheless, at least where the uncertainty in signal measurements is low (see below), we can predict the orientation from the scattering by utilising computational techniques such as neural networks. We thus utilised the Python machine learning program \textit{scikit-learn} to build a neural network for identifying the dimer's orientation from its light scattering signal. The network was trained by generating a database of random orientation vectors, calculating the corresponding light scattering signals, and then using the network to estimate the probability of a given signal coming from a dimer in a given reference orientation. The network's loss function was evaluated and used to improve the estimation, the network being trained until the improvement in the loss function was less than 0.0001. 
%
Importantly, the estimation provided by the neural network can be improved further by accounting for any prior information we know about the dimer,  utilising Bayesian inference to update the neural network's estimation:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Calculation of error}
\label{sec:divergence}
To evaluate the above estimation of dimer orientation from scattering signal, we use a Brownian simulation of a dimer in the optical trap (Section~\ref{sec:brownian}) to compare estimated most probable reference orientation, derived from the dimer's scattering through Eq.~\eqref{eq:bayes}, with the dimer's known \emph{actual} orientation $\hat{\bf s}$. MSTM provides calculated light scattering from the simulated dimer $I(\hat{\bf s}, \theta)$ and we use \eqref{eq:scale} to obtain normalized values at each measurement angle $\theta_k$,  $y_1(\hat{\bf s})$, $y_2(\hat{\bf s})$, $y_3(\hat{\bf s})$, from which we obtain $p(\hat{\bf n}_{\alpha}\parallel y_1, y_2, y_3)$. Because we know the actual orientation $\hat{s}$ we can measure the error in the model's estimate by comparing the reference orientation closest to $\hat{s}$, denoted as $\hat{\bf n}_{best}$, with the most probable predicted orientation from Eq.~\eqref{eq:bayes}. An ideal result would be one where the probability distribution is 0 for every $\hat{\bf n}$ apart from $\hat{\bf n}_{best}$:
\begin{align}
	\label{eq:best}
	p_{best} = 
	\begin{cases}
		1 & \text{when $\hat{\bf n}_\alpha$ = $\hat{\bf n}_{best}$}\\
		0 & \text{anywhere else}
	\end{cases}
\end{align}
In reality the distribution from Eq.~\eqref{eq:bayes} will assign some non-zero
probability to every reference orientation, leading to some level 'confidence' in orientation prediction, which can be quantified by calculating the Kullback-Leibler divergence $K_l$ between the two distributions:
\begin{align}
	K_{l, \#}(p_{best}\parallel p(\hat{\bf n}_\alpha| y_1, y_2, y_3))
	= p_{best}\ln \left[\frac{p_{best}}{p(\hat{\bf n}_{best}| y_1,y_2,y_3))}
	\right]
	\label{eq;kullback}
\end{align}
where a larger value of $K_l$ indicates that our model is less confident in its
prediction of the dimer's orientation. The divergence $K_l$ thus illustrates the 'spread'
in the estimated dimer orientation probability --- a distribution strongly peaked at 
some value would give us more confidence in that value than a near-uniform distribution 
where the scattering measurement could imply a wide range of possible orientations --- 
but it does not directly indicate our estimate's actual accuracy. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and  Discussion}
\label{sec:Discussion}
\subsection{Minimal number of detectors}
\label{sec:detectors}
The exact number of detectors was initially assumed to be arbitrary, in that it made no difference to our estimate whether we used 2 angles or 200. For practical purposes it seemed beneficial that we demonstrate our method works for a minimal number of detection angles, as geometric constraints come into play when trying to install a high number of detection fibres around a single point.

Initially we tried capturing scattered light only in the x-y plane as to simulate the detectors all being placed on a glass cover slip. The results of this estimate are shown below. As can be seen from Figure~ the neural network performs particularly poorly in this situation, seemingly jumping back and forth between different reference orientation at random. We found that when all three detectors are in the same plane the expected signal can appear identical despite the dimer being in completely different orientations. This is shown in Figure~ which plots the expected signals from 30 reference orientations, each point is labelled with its corresponding reference orientation, the fact that points have multiple labels shows that the dimer's scattering is indistinguishable in these two reference orientation. More specifically, if the detectors are placed say in the x-y plane then only when the dimer is pointed nearly fully upright will the expected signal be entirely unique. This is illustrative of the difficulty behind the inverse light scattering problem; as attributing the measured signal will not reveal any new information about the entity unless we can remove uncertainty around other aspects of the scatterer.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.49\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/fig2b.png}
	\end{subfigure}
	\begin{subfigure}{0.4\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/intensity_space.png}
	\end{subfigure}
	\caption{(a) 30 Reference orientations represented by vectors pointing from [0,0,0] to each point (b) Expected signal for each reference orientation, some signals map to multiple reference orientation and are labelled as such.}
	\label{fig:angles}
\end{figure}

To remedy this we raise the third detector out of the x-y plane; as such the expected signals from each reference orientation is unique. With only three detectors the difference in expected signals can appear insignificant; by adding a $4^{th}$ detector we can differentiate signals more reliably, improving the neural networks performance. In line with our goal of making this method viable in a laboratory setting we decided not to increase the number of detectors further than 4. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing the Model}
\label{sec:test}
Using our simulation from Section~\ref{sec:brownian} we simulated the motion of a silica dimer ($n = 1.59$) trapped in water ($n = 1.33$) within a 5 mW optical trap. The trapping laser is 1064nm NIR focused through a 1.25 NA objective. The dimer is comprised of two tangent spheres with radii $1 \mu m$ and $0.5 \mu m$ respectively. We simulated the first 10 seconds of motion, calculating the orientation and position every 1 ms.

@@

One reason we observe such large jumps in orientation estimated from scattering signals is that there is no simple correlation between the 'distance in scattering space' between scattering signals from two different orientations, and their separation in orientation space: even a large change in orientation can involve a small change in scattering. Combining this fact with use of a uniform prior, indicating essentially no knowledge of how orientation should behave, there is no constraint on how much estimated orientation can change from time-step to time-step. To improve the estimation we can therefore use knowledge of the physical limitations of the object in the trap and its dynamics, imposing a more physically grounded prior, accounting in this case for the fact that the motion of the dimer is limited due to the trap stiffness. Here the prior of the reference orientations $p(\hat{\bf n}_\alpha)$ was redefined at each time step as a Boltzmann distribution of the physical distance between the previous estimate $\hat{\bf n}_{est}(t-\Delta t)$ and each reference orientation $\hat{\bf n}_\alpha$. Put simply, we are reweighing our estimation based on the size of rotation required, with smaller movements being favoured over large movements:

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Asymmetric dimer dynamics}
%\label{sec:motion}

%The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\,\mu{\rm m}$, $a_2=0.5\,\mu{\rm m}$) over the first $10$ seconds of entering the optical trap.  The initial orientation was assumed as strictly vertical (in line with the beam propagation direction). The dimer's position and orientation was recorded every $1\,{\rm ms}$ for using as a test dataset for our model.

%\begin{figure}[h]
%	\centering
%	\begin{subfigure}{0.45\textwidth}
	%	\subcaption{}
	%	\includegraphics[width =\textwidth]{./Images/fig5a.png}
%	\end{subfigure}
%	\begin{subfigure}{0.45\textwidth}
%		\subcaption{}
%		\includegraphics[width=\textwidth]{./Images/fig5b.png}
%	\end{subfigure}
%	\caption{Simulation results of: (a) the dimer's orientation vector with time, (b) the dimer's [x,y,z] position with time.}
%	\label{fig:motion}
%\end{figure}

%As can be seen from Figure~\ref{fig:motion}, the dimer undergoes a full $180^{\circ}$ rotation upon entering the trap.  Typically horizontal alignment of a dimer is unstable and will result in the particle rotating to align along its vertical axis.  It is interesting to note that the dimer is furthest from the trap centre as it reaches a horizontal orientation before drawing closer again as it inverts completely. Further simulations of dimers with different size ratio $a_1 \geq 2a_2$ showed a similar behaviour of initially moving away from the trap focus and rotating while remaining trapped before re-aligning, while dimers with size ratios $a_1 < 2a_2$ immediately aligned into a fixed vertical position.
%
%In the simulations of Vigilante~\emph{et~al}.\ \cite{Vigilante2020Brownian_OT}, trapped symmetrical dimers were investigated; their findings showed that the optical torque on the dimer goes to zero while aligned vertically and is at its maximum in a horizontal alignment. Therefore, the inversion of an asymmetric dimer suggests that if the size difference is significant the optical torque is minimal for a dimer in both horizontal and vertical orientations.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Impact of measurement noise on model predictions}
\label{sec:epsilon}

So far a key assumption of the neural network implementation is that the detected scattering signal has no uncertainty associated with it. In reality of course scattering signals will always have some non-zero measurement noise. This can be attributed to a variety of factors, from a measurement bias in the detector, to the Brownian motion of the dimer itself. To explore the impact of measurement uncertainty on orientation estimation model performance we introduce a Gaussian noise to the measured signal:
\begin{align}
	I(\hat{\bf s}) = I(\hat{\bf s}) \pm \epsilon I(\hat{\bf s})
\end{align}
where $\epsilon$ is the percentage error associated with the scattering signal.  Figure~\ref{fig:epsilon} shows the performance of the model at a range of $\epsilon$ using in-plane detector angles $15^{\circ}$, $55^{\circ}$, $90^{\circ}$ and out-of-plane detector at $75^{\circ}$, with $\beta$ set to $1$:
